HSF Community White Paper on Data Organization, Management and Access (DOMA)

### 2018-03-25 CWP/papers/HSF-CWP-2017-04_doma/latex/hsf-cwp-dm.tex

\subsection{Sub-file granularity}\label{sub-file-granularity}
Sub-file granularity ({\it e.g.} event-based granularity) should be studied to see
whether it can be implemented efficiently and in a scalable, cost-effective manner for
all use cases that rely on event selection and whether it offers an advantage over
the current paradigm of file-based granularity.

\begin{itemize}
\item Quantify impact on performances
  and resource utilization (storage, network) for the main type of
  access patterns (simulation, reconstruction, analysis)
\item Assess impact on catalogs and
  data distribution
\item Assess whether
  event-granularity makes sense in object stores that tend to require
  large chunks of data for efficiency
\item Does it improve recoverability
  from preemption, in particular when using cloud spot resources and/or
  dynamic HPC resources
\end{itemize}

@@

\subsection{Data organization and analysis
technologies}\label{data-organization-and-analysis}
Data organization and analysis technologies in use by other big data users should be
studied to glean whether we can benefit from them.

\begin{itemize}
\item Row-based vs. column-based
  organization: impact of storage organization on the performance of
  each kind of access, potential storage format providing good
  performance for both
\item Can map-reduce, Spark-like
  analysis and their functional or declarative interfaces be adapted to
  HEP analysis needs?
\item Evaluate just in time
  decompression schemes and mappings onto hardware architectures
  considering the flow of data from spinning disk to memory and
  application
\end{itemize}

@@

\subsection{Data caching}\label{data-caching}
Investigate the role that data placement optimizations, such as caching,
can play in optimizing the use of computing resources as well as the
technologies that can be used for this.

\begin{itemize}
\item Quantify benefit of caching for
  main use cases (reconstruction, analysis, simulation)
\item Assess benefit of caching for
  Machine Learning-based applications, in particular for the learning
  phase
\item What are the benefits of a CDN
  or NDN approach beyond what HEP is doing today with their data
  delivery systems?
\end{itemize}

\emph{The first two actions should be completed by 2020 but CDN/NDN
evaluation is more long-term}

\subsection{Getting the most out of the
storage diversity}\label{storage-diversity}
How to minimize HEP infrastructure cost by
exploiting varied quality of service from different storage
technologies?

\begin{itemize}
\item Which role for
  opportunistic/tactical storage
\item Role of archival storage
  solutions
\end{itemize}

\emph{Proof-of-concept by 2020, implementation if PoC successful in the following years}

\subsection{Efficiency / latency (global
optimization}\label{efficiency-latency}
Can we globally optimize data access latency vs
efficiency of CPU use at a sustainable cost?

\begin{itemize}
\item Impact of concentrating the
  data in fewer, larger locations (``data-lake'' approach)
\item Impact of an increased use of
  opportunistic compute resources, further from the data
\end{itemize}

\emph{Proof-of-concept by 2020, implementation if PoC successful in the following years}

>>>>

\subsection{Sub-file granularity}\label{sub-file-granularity}
Sub-file granularity ({\it e.g.} event-based granularity) should be studied
to see whether it can be implemented efficiently and in a scalable, cost-effective
manner for all use cases that rely on event selection and whether it offers an advantage
over the current paradigm of file-based granularity. Proposed actions are:

\begin{itemize}
\item Quantify impact on performances
  and resource utilization (storage, network) for the main type of
  access patterns ({\it i.e.} simulation, reconstruction, analysis).
\item Assess the impact on catalogs and
  data distribution.
\item Assess whether
  event-granularity makes sense in object stores that tend to require
  large chunks of data for efficiency.
\item Test for improvement in recoverability
  from preemption, in particular when using cloud spot resources and/or
  dynamic HPC resources.
\end{itemize}

@@

\subsection{Data organization and analysis
technologies}\label{data-organization-and-analysis}
Data organization and analysis technologies in use by other big data users should be studied
to glean whether we can benefit from them. Proposed actions are:

\begin{itemize}
\item Evaluate row-based versus column-based
  data organization on the performance of each type of access used.
\item Investigate data storage and access solutions that support the use of map-reduce or
Spark-like analysis tools and their adaptability to HEP analysis needs.
\item Evaluate just-in-time
  decompression schemes and mappings onto hardware architectures
  considering the flow of data from spinning disk to memory and
  application.
\end{itemize}

@@

\subsection{Data caching}\label{data-caching}
Investigate the role that data placement optimizations, such as caching, can play in optimizing
the use of computing resources as well as the technologies that can be used for this. Proposed actions are:

\begin{itemize}
\item Quantify the benefit of caching for reconstruction, analysis, and simulation.
\item Assess the benefit of caching for
  Machine Learning-based applications, in particular for the learning
  phase.
\item Evaluate the benefits of using different approaches to data delivery from what is currently in use in HEP.
Namely, Content Delivery Networks (CDN) and Named Data Networking (NDN) should be studied.
\end{itemize}

-\emph{The first two actions should be completed by 2020 but CDN/NDN
-evaluation is more long-term}

\subsection{Getting the most out of the
storage diversity}\label{storage-diversity}

The landscape of increasing diversity in storage service offerings and technologies should be studied and
exploited in order to reduce HEP infrastructure costs. The proof-of-concept phase involves the following tasks:

\begin{itemize}
\item Understand what role tactical or opportunistic (potentially short-term) storage can play in HEP.
\item Re-evaluate the role of archival storage solutions for HEP.
\end{itemize}

\emph{The  proof-of-concept should be completed by 2020 and, if successful, implementation in
production will be done in the following years.}

\subsection{Global optimization of efficiency and latency}\label{efficiency-latency}
The inherent trade-offs between data access latency and efficiency of CPU use need to be studied and optimized
on a global scale across the HEP infrastructure. The proof-of-concept phase involves the following tasks::

\begin{itemize}
\item Understand the impact of concentrating the
  data in fewer, larger locations (the ``data-lake'' approach).
\item Understand the impact of an increased use of
  opportunistic compute resources, specifically those further from where data are stored.
\end{itemize}

\emph{The  proof-of-concept should be completed by 2020 and, if successful, implementation in
production will be done in the following years.}
